---
title: "Week 7 Homework"
author: "Cole Jower"
date: "October 4, 2017"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Upload File
```{r upload file}

df= read.csv("../data/california-state-university-2015.csv")

```

## Convert file to characters
``` {r convert to characters, results = "hide"}

df$Employee.Name = as.character(df$Employee.Name)

```

## Find all entries with name 'Martinez'
```{r Find all entries with name "Martinez", results = "hide"}

df$Employee.Name[grepl(pattern = "Martinez", df$Employee.Name)]

```

## Everyone whose names starts with the letters 'Wil*'
``` {r Everyone whose name starts with the letters "Wil*", results = "hide"}

df$Employee.Name[grepl(pattern = "Wil", df$Employee.Name)]
```

## Everyone with a hyphenated name
``` {r Everyone with a hyphenated name, results = "hide"}

df$Employee.Name[grepl(pattern = "[[:alpha:]]-[[:alpha:]]", df$Employee.Name)]

```

## Everyone whose middle name is Paul
``` {r Everyone whose middle name is Paul, results = "hide"}

df$Employee.Name[grepl(pattern = "[[:blank:]]Paul[[:blank:]][[:alpha:]]", df$Employee.Name)]

```

## All people with the last name Martinez
``` {r All people with the last name Martinez, results = "hide"}

df$Employee.Name[grepl(pattern = "[[:alpha:]][[:blank:]]Martinez", df$Employee.Name)]

```

## All people with last name Martinez (but no hyphenated names)
``` {r All people with last name Martinez (but no hyphenated names), results = "hide"}

df$Employee.Name[grepl(pattern = "[[:blank:]]Martinez$", df$Employee.Name)]

```

## First name starts with E, has a middle initial, and last name is Martinez
``` {r First name starts with E, has a middle initial, and last name is Martinez}

ENames = df$Employee.Name[grepl(pattern = "^E", df$Employee.Name)]
ENames[grepl(pattern = "[[:blank:]][[:alpha:]][[:blank:]]Martinez", ENames)]

```

## First name starts with E, has no middle initial, and last name is Martinez
``` {r First name starts with E, has no middle initial, and last name is Martinez}

df$Employee.Name[grepl(pattern = "^\\<E[a-z]*.\\sMartinez$", df$Employee.Name)]


```