---
title: "Project Portfolio Outline"
author: "Cole Jower"
date: "October 30, 2017"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### For my portfolio, I will be using data from my work (San Francisco Bay Bird Observatory [SFBBO]).  This data was gathered using the Daubenmire cover class method at a restoration site in Alviso, known as pond A17.  This pond was formerly an industrial salt evaporation pond that was obtained by the Don Edwards National Wildlife Refuge, a part of the US Fish and Wildlife Service, and underwent a levee breach to return it tidal marsh.  As a part of the SFBBO habitat restoration program, we seeded and transplanted several species of native upland and transitional tidal marsh species in 30 mounds placed on old levee, and have been monitoring their recruitment at the site for nearly a year.  For my portfolio, my aim is to transpose the data in its current format to a long list format, and then analyze it to determine restoration success.  Part of success will be to determine if we have met refuge goals of continuous vegetation cover at a vegetation height of 30cm and greater.  In addition we want to determine which of our methods (if any) are more suitable for restoration.  In this regard we will look at different soil treatments at the restoration site, with 30 mounds that were either treated with straw or raking, with various types of soil including hybrid, compost, or organic.

### Install new packages to transform data
``` {r install reshape2 package}
install.packages("reshape2", dependencies = T)
```

##### MANIPULATE DATA ######

## set up datasheet
```{r}
library(reshape2)

data.wide <- read.csv(file.choose(), stringsAsFactors = F, strip.white = T, na.strings = c(""))
# HabitatsDBCurrentFormat.csv

head(data.wide)
```

## extract just the top 5 lines
``` {r extract top 5 lines}

heading <- data.wide[1:5,c(1,7:nrow(data.wide))]

```

## transpose heading
``` {r transpose heading}

heading.t <- as.data.frame(t(heading), stringsAsFactors = F)
cnames <- as.character(unlist(heading.t[1,]))
heading.t <- heading.t[-1,]
colnames(heading.t) <- cnames

```

## extract data from row names
``` { r extract from rows}

heading.t$YYYYMMDD_TRANSECT_QUAD <- rownames(heading.t)
heading.t$Date <- sapply(strsplit(heading.t$YYYYMMDD_TRANSECT_QUAD, split = "_", fixed = T),"[",1)
heading.t$TransectCode <- sapply(strsplit(heading.t$YYYYMMDD_TRANSECT_QUAD, split = "_", fixed = T),"[",2)
heading.t$QuadratID <- sapply(strsplit(heading.t$YYYYMMDD_TRANSECT_QUAD, split = "_", fixed = T),"[",3)

heading.t$Date <- gsub("X", "", heading.t$Date, fixed = T)
heading.t$Date <- as.Date(heading.t$Date, "%Y%m%d")

heading.t <- heading.t[,-which(colnames(heading.t)=="Scientific Name")]

```

## prepare the rest of the data sheet 
``` {r prepare the rest}

body <- data.wide[6:nrow(data.wide),]
colnames(body)[1:6] <- as.character(unlist(data.wide[5,1:6]))

```

## remove columns that are entirely full of NA's
```{r remove NA columns}

body <- body[, unlist(lapply(body, function(x) !all(is.na(x))))]

body.t <- melt(body, id = c("Scientific Name","Common Name","Code","Status","Grass/Forb","Annual/Perennial"), variable.name = "YYYYMMDD_TRANSECT_QUAD")
body.t$Presence <- substr(body.t$value,1,1)
body.t$HeightClass <- substr(body.t$value,2,2)

```

## merge the heading part with the body part
```{r merge heading with body}

data.all <- merge(heading.t, body.t[,-which(colnames(body.t)=="value")], by = "YYYYMMDD_TRANSECT_QUAD", all = T)

```

## export the columns of interest to a csv
```{r export as csv}

write.csv(data.all, file = "Habitats_Transposed.csv") 

```

## compare soil treatments
``` {r}
### Insert script here

```

## examine cover class and heights
```{r}
### Insert script here
### End product will be a histogram showing relative frequency of species occurence (y axis) by height class (x axis)
### Another histogram will examine relative frequency by height class, but only with species that fell within the require cover class 
```
```