---
title: "Portfolio, Update 3.0"
author: "Cole Jower"
date: "November 29, 2017"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Portfolio Description
For my portfolio, I will be using data from my work (San Francisco Bay Bird Observatory [SFBBO]).  This data was gathered using the Daubenmire cover class method at a restoration site in Alviso, known as pond A17.  This pond was formerly an industrial salt evaporation pond, obtained by the Don Edwards National Wildlife Refuge, a part of the US Fish and Wildlife Service, and underwent a levee breach to return it tidal marsh.  As a part of the SFBBO habitat restoration program, we seeded and transplanted several species of native upland and transitional tidal marsh species in 30 mounds placed on old levee, and have been monitoring their recruitment at the site for nearly a year.  For my portfolio, my aim is to transpose the data in its current format to a long list format, and then analyze it to determine restoration success.  For this project, I will only be examining the month of June (for 2017), a time deemed appropriate due to the high presence of an invasive weed, slenderleaf iceplant (Mesembryanthemum nodiflorum). Thus, native plant performance (determined by relative frequency) will be examined in the presence of a persistent and invasive weed. Success will be to determine if we have met refuge goals of continuous vegetation cover (common presence) at a vegetation height of 30cm and greater.  In addition we want to determine which of our methods (if any) are more suitable for restoration.  In this regard we will look at different soil treatments at the restoration site, with 30 mounds that were trop dressed with compost, hybrid, or organic soils, at varying depths (1, 2, or 3 inches).  In summary, my output will contain three sets of histograms with the following information:

1) Compare Native and Weed presence by height class (which is structurally more dominant?)

2) Examine Native and Weed canopy presence by heights class (are present Natives reaching refuge goals?)

3) Examine native species relative frequency relative to soil depth and material (do certain treatments work better than others?)

## Transpose Data
``` {r TransposeData, results = "hide"}

# Install Packages
install.packages("reshape2", dependencies = T, repos = "http://cran.us.r-project.org") # only have to do this once


# Manipulate Data
library(reshape2)

data.wide <- read.csv(file.choose(), stringsAsFactors = F, strip.white = T, na.strings = c(""))
# A17JuneData.csv
head(data.wide)

# extract just the top 9 lines
heading <- data.wide[1:9,c(1,7:nrow(data.wide))]

# transpose heading
heading.t <- as.data.frame(t(heading), stringsAsFactors = F)
cnames <- as.character(unlist(heading.t[1,]))
heading.t <- heading.t[-1,]
colnames(heading.t) <- cnames

# extract data from row names
heading.t$YYYYMMDD_TRANSECT_QUAD <- rownames(heading.t)
heading.t$Date <- sapply(strsplit(heading.t$YYYYMMDD_TRANSECT_QUAD, split = "_", fixed = T),"[",1)
heading.t$TransectCode <- sapply(strsplit(heading.t$YYYYMMDD_TRANSECT_QUAD, split = "_", fixed = T),"[",2)
heading.t$QuadratID <- sapply(strsplit(heading.t$YYYYMMDD_TRANSECT_QUAD, split = "_", fixed = T),"[",3)

heading.t$Date <- gsub("X", "", heading.t$Date, fixed = T)
heading.t$Date <- as.Date(heading.t$Date, "%Y%m%d")

heading.t <- heading.t[,-which(colnames(heading.t)=="Scientific Name")]

# prepare the rest of the data sheet (other than the heading)
body <- data.wide[10:nrow(data.wide),]
colnames(body)[1:6] <- as.character(unlist(data.wide[9,1:6]))

# remove columns that are entirely full of NA's
body.t <- body[, unlist(lapply(body, function(x) !all(is.na(x))))]

body.t <- melt(body, id = c("Scientific Name","Common Name","Code","Status","Grass/Forb","Annual/Perennial"), variable.name = "YYYYMMDD_TRANSECT_QUAD")
body.t$Presence <- substr(body.t$value,1,1)
body.t$HeightClass <- substr(body.t$value,2,2)

# merge the heading part with the body part
data.all <- merge(heading.t, body.t[,-which(colnames(body.t)=="value")], by = "YYYYMMDD_TRANSECT_QUAD", all = T)

# subset out remaing NAs to only get rows with data
final.data <- subset(data.all, Presence!="NA" )

final.data <- subset(final.data, select = -YYYYMMDD_TRANSECT_QUAD)

# export the columns of interest to a csv
write.csv(final.data, file = "A17June_Transposed.csv") 
# This final format will be imported to a Microsoft Access database

```

```{r read.csv}
# Open .csv
A17 <- read.csv('../Portfolio/A17June_Transposed.csv')
```

# 1) Compare Native and Weed presence by Height Class
``` {r NativeVsWeedHeightClass}

# Create contingency table comparing status (native or weed) versus height class
StatusVSHeight <- table(A17$Status, A17$HeightClass)
NativeVsWeed<- as.data.frame.matrix(StatusVSHeight)

# Calculate relative frequency
RelFreqNVW <-(NativeVsWeed/nrow(A17))*100

# Convert table to numeric
NatVsWeedHeight <- data.matrix (RelFreqNVW)

# Create side-by-side relative frequency graph
barplot(NatVsWeedHeight, ylab = "Relative Frequency", xlab = "Height Class", main = "Native/Weed Presence by Height Class", col=c("lightblue", "Orange"),names.arg=c("1 (<30cm)", "2 (30-45cm)","3 (45-60cm)","4 (>60cm)"), beside = TRUE, width = .3, legend = rownames(NatVsWeedHeight), cex.names = 0.75)


```

# 2) Examine Native and Weed canopy presence by heights class
``` {r CommonPresencebyHeights}

# Subset all presence by Common
C.Presence <- subset(A17, Presence == "C")
TableC <- table(C.Presence$Status, C.Presence$HeightClass)
Common <- as.data.frame.matrix(TableC)

# Calculate Relative Frequency
CRF <- (Common/ nrow(C.Presence)*100)

# Convert table to numeric
CRF.Num <- data.matrix (CRF)

# Subset all presence by Uncommon
U.Presence<- subset(A17, Presence == "U")
TableU <- table(U.Presence$Status, U.Presence$HeightClass)
Uncommon <- as.data.frame.matrix(TableU)

# Calculate Relative Frequency
URF <- (Uncommon/ nrow(U.Presence)*100)

# Convert table to numeric
URF.Num <- data.matrix (URF)

# Subset all presence by Rare
R.Presence<- subset(A17, Presence == "R")
TableR<- table(R.Presence$Status, R.Presence$HeightClass)
Rare <- as.data.frame.matrix(TableR)

# Calculate Relative Frequency
RRF <- (Rare/ nrow(R.Presence)*100)

# Convert table to numeric
RRF.Num <- data.matrix (RRF)

# Create histograms
barplot(CRF.Num, ylab = "Relative Frequency", xlab = "Height Class", main = "Common Native/ Weed Presence by Height Class", col=c("Green", "Violet"), beside = TRUE, names.arg=c("1 (<30cm)", "2 (30-45cm)","3 (45-60cm)","4 (>60cm)"),legend = rownames(CRF.Num), ylim = c(0,50), cex.names = .75)

barplot(URF.Num, ylab = "Relative Frequency", xlab = "Height Class", main = "Uncommon Native/ Weed Presence by Height Class", col=c("Green", "Violet"), beside = TRUE, names.arg=c("1 (<30cm)", "2 (30-45cm)","3 (45-60cm)"),legend = rownames(URF.Num), ylim = c(0,60))

barplot(RRF.Num, ylab = "Relative Frequency", xlab = "Height Class", main = "Rare Native/ Weed Presence by Height Class",col=c("Green", "Violet"), beside = TRUE, names.arg=c("1 (<30cm)", "2 (30-45cm)","3 (45-60cm)"),legend = rownames(RRF.Num), ylim = c(0,80))

```

# 3) Examine native species relative frequency relative to soil depth and material
``` {r Depth&MaterialVSCoverClass}

# Subset native species
Natives = A17[A17$Status == "Native",]

# Native presence by Compost, 1 inch
NativeCompost = Natives[Natives$Material == "Compost",]
NativeCompost1 = NativeCompost[NativeCompost$Soil.Depth == "1",]

NC1 <- table(NativeCompost1$Status, NativeCompost1$Presence)
NC1Table <- as.data.frame.matrix(NC1[,c(2,3,1)])

NC1Freq <- (NC1Table/nrow(NativeCompost1))*100
NC1RelFreq <-as.matrix(NC1Freq)

# Native presence by Compost, 2 inch
NativeCompost2 = NativeCompost[NativeCompost$Soil.Depth == "2",]

NC2 <- table(NativeCompost2$Status, NativeCompost2$Presence)
NC2Table <- as.data.frame.matrix(NC2[,c(2,3,1)])

NC2Freq <- (NC2Table/nrow(NativeCompost2))*100
NC2RelFreq <- as.matrix(NC2Freq)

# Native presence by Compost, 3 inch
NativeCompost3 = NativeCompost[NativeCompost$Soil.Depth == "3",]

NC3 <- table(NativeCompost3$Status, NativeCompost3$Presence)
NC3Table <- as.data.frame.matrix(NC3[,c(2,3,1)])

NC3Freq <- (NC3Table/nrow(NativeCompost3))*100
NC3RelFreq <- as.matrix(NC3Freq)

# Combine all compost graphs into one 
NCALL = rbind(NC1RelFreq, NC2RelFreq, NC3RelFreq)

###################################################################################


# Native presence by Hybird, 1 inch
NativeHybrid = Natives[Natives$Material == "Hybrid",]
NativeHybrid1 = NativeHybrid[NativeHybrid$Soil.Depth == "1",]

NH1 <- table(NativeHybrid1$Status, NativeHybrid1$Presence)
NH1Table <- as.data.frame.matrix(NH1[,c(2,3,1)])

NH1Freq <- (NH1Table/nrow(NativeHybrid1))*100
NH1RelFreq <-as.matrix(NH1Freq)

# Native presence by Hybrid, 2 inch
NativeHybrid2 = NativeHybrid[NativeHybrid$Soil.Depth == "2",]

NH2 <- table(NativeHybrid2$Status, NativeHybrid2$Presence)
NH2Table <- as.data.frame.matrix(NH2[,c(2,3,1)])

NH2Freq <- (NH2Table/nrow(NativeHybrid2))*100
NH2RelFreq <- as.matrix(NH2Freq)

# Native presence by Hybrid, 3 inch
NativeHybrid3 = NativeHybrid[NativeHybrid$Soil.Depth == "3",]

NH3 <- table(NativeHybrid3$Status, NativeHybrid3$Presence)
NH3Table <- as.data.frame.matrix(NH3[,c(2,3,1)])

NH3Freq <- (NH3Table/nrow(NativeHybrid3))*100
NH3RelFreq <- as.matrix(NH3Freq)

# Combine all Hybrid graphs into one 
NHALL = rbind(NH1RelFreq, NH2RelFreq, NH3RelFreq)

###################################################################################

# Native Organic, 1 inch
NativeOrganic = Natives[Natives$Material == "Organic",]
NativeOrganic1 = NativeOrganic[NativeOrganic$Soil.Depth == "1",]

NO1 <- table(NativeOrganic1$Status, NativeOrganic1$Presence)
NO1Table <- as.data.frame.matrix(NO1[,c(2,3,1)])

NO1Freq <- (NO1Table/nrow(NativeOrganic1))*100
NO1RelFreq <-as.matrix(NO1Freq)

# Native presence by Organic, 2 inch
NativeOrganic2 = NativeOrganic[NativeOrganic$Soil.Depth == "2",]

NO2 <- table(NativeOrganic2$Status, NativeOrganic2$Presence)
NO2Table <- as.data.frame.matrix(NO2[,c(2,3,1)])

NO2Freq <- (NO2Table/nrow(NativeOrganic2))*100
NO2RelFreq <- as.matrix(NO2Freq)

# Native presence by Organic, 3 inch
NativeOrganic3 = NativeOrganic[NativeOrganic$Soil.Depth == "3",]

NO3 <- table(NativeOrganic3$Status, NativeOrganic3$Presence)
NO3Table <- as.data.frame.matrix(NO3[,c(2,3,1)])

NO3Freq <- (NO3Table/nrow(NativeOrganic3))*100
NO3RelFreq <- as.matrix(NO3Freq)

# Combine all organic graphs into one 
NOALL = rbind(NO1RelFreq, NO2RelFreq, NO3RelFreq)

# Create histograms
par(mfrow=c(3,1),mar = c(2, 2, 2, 2) + 0.1) 
barplot(NCALL, ylab = "Relative Frequency", main = "Compost, Native Species by Cover Class", col = c("blue", "green", "red"), names.arg=c("Rare (0-5%)","Uncommon (5-33%)","Common (33-100%)"), beside = TRUE, width = .3, ylim = c(0,60))
barplot(NHALL, ylab = "Relative Frequency", main = "Hybrid, Native Species by Cover Class", col = c("blue", "green", "red"), names.arg=c("Rare (0-5%)","Uncommon (5-33%)","Common (33-100%)"), beside = TRUE, width = .3, ylim = c(0,60))
barplot(NOALL, ylab = "Relative Frequency", xlab = "Cover Class", main = "Organic, Native Species by Cover Class", col = c("blue", "green", "red"), names.arg=c("Rare (0-5%)","Uncommon (5-33%)","Common (33-100%)"), beside = TRUE, width = .3)

### Need to add legend for soil depths ###
```

#### One meter look around against height for all months, as well as september
